Asymmetric relationships and their effects on coexistence

### 2023-06-30 paper/RefereeResponses1.tex

\noindent ***\emph{<Address the issue of whether and to what extent the process of removing ATAs is unique – probably just with a one-sentence callout in the main text, and some material in the SM (possibly a new section)>} \\

\noindent C6) The notation by which the partial sharp variables are introduced was a little unintuitive to me.  For example, when the authors compare the expectation values 
$E[r(E,C)]$ and $E[r(E^{||},C^{||})]$, what is really changing is the distribution over which the expectation values are evaluated.  Maybe this is standard notation, but somehow it may be clearer to put the subscript on the E for expectation value, to indicate that the probability distribution used to generate the definition of the averaging is what differs between the two expressions, while the variables E and C themselves are in the end just dummy variables to be integrated over.  \\

\noindent ***\emph{<This one asks for possibly changing notation. I don’t think we should do it, but we need to explain in a response what is actually going on, why we chose the notation that we do, and then add a new SM section addressing what the referee is talking about. Actually it’s the bivariate distribution that changes when you apply ||, tho the marginals are not changed. So the referee does have a bit of a point – E^{||} by itself is actually the same distribution as E, it’s just that it’s a different random variable in the sense of being related differently to C^{||} than E is to C. This is the same issue with the # notation that has already been established. It’s actually formally correct in the sense of random variables if we view *all* our random variables as being measurable functions from the *same* probability space. So in the end the referee does not really have a point if one wants to be formal. But you can explain all this in a new sup mat section.>} \\

\noindent C7) It was not clear to me what the reasoning was for including the correlation and ATAs in abundance for the two planktonic species in Figure 1(d) and (e).  These ATAs in abundances don't in an obvious (to me) way translate into ATAs of the underlying environment and competition variables for each species. Is this really a demonstration that ATAs occur for C and E?  If not, then it seems to be like saying ATAs exist somewhere in nature, and therefore might be present in our focal situation, which is not such a very strong reason to show (d) and (e).  Maybe I am missing something. \\

@@

\noindent ***\emph{<Yes, the referee is right about this, that's what we're doing. Just admit to it and justify doing it anyway because a) we were asked to include this kinda thing by poeple who read it before we submitted; b) the value is not only in showing that ATAs occur, but also in helping people understand more fully what they look like; and c) we cannot show ATAs in E and C (at least not this early in the paper), becuase no one has looked at this before so there is no prior evidence of it.>} \\

\noindent C8) The authors mention that $r_{i \bs i}$ indicates growth rate of species when $i$ is rare and $j$ is at steady state.  I think $r_{j \bs i}$ must also assume that $j$ is at steady state, but this isn't mentioned. Perhaps that could be added for completeness. \\

@@

\noindent C9) E.g. line 100 needs removal of the word "are" or the addition of the word "and" somewhere to make sense. I'd consider going through the paper carefully for any typos or small corrections of this kind. \\

\noindent ***\emph{response} \\

\noindent \textbf{Comments of referee 2:} \\

\noindent C10) In their paper the authors decompose the “storage effect” into two separate effects, what they call the “asymmetric tail association” and the covariance term per se. The authors explain well why they do this and why they think this may be important. Additionally, they show the best possible efforts to convince the readers why such a decomposition may be important and what we may learn from it. Overall, the paper is interesting and quite understandable. I think the authors have done a very good job. \\

@@

\noindent ***\emph{Thank you for the very positive feedback!} \\

\noindent C11) Unfortunately, I’m personally not convinced that storage effect (and therefore probably ATA) are relevant for coexistence in nature and therefore I’m sceptical about this new approach. I want to be clear that this is, in part, my personal opinion and I know experts in the field which have a very different view. What the authors have shown (and it appears to be correct) is that in a model where niche differentiation via different resources and predators is impossible, then ATA can be an important driver of coexistence, and I agree. However, in every application where both storage effect, relative non-linearity and niche differentiation have been measured, there we see that storage effect and relative non-linearity where very small compared to simple resource differentiation. But, one might argue, that we have barely any such empirical applications, and I would agree, which is why I label this my personal opinion. (The few papers that I know of are \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1002%2Fecy.2726&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=HbUTjzCNQrvGSARY5u3Eqme88DbKiNyiAVNOUGzjcbE%3D&reserved=0}{Zepeda et al. 2019, Ecology}, \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1890%2F14-1741.1&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=KKIBuLjBN2vNGWfEpduAGGrMAt7ov9a4F53CmlZQlO0%3D&reserved=0}{Chu and Adler 2015}, and Ellner et al. 2019, see also the new paper by \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fesajournals.onlinelibrary.wiley.com%2Fdoi%2Fepdf%2F10.1002%2Fecm.1585&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=YaqsZ6axu6rYV12kh9IiRIbES7ZUMWIN7HXgB08iPzc%3D&reserved=0}{Stump and Vasseur}, which I have not yet read fully) In my understanding, ATA is a decomposition of storage effect into even more components, therefore I would expect, on average the ATA effect to be slightly smaller than storage effect (but, as you have shown can also be larger than storage effect). I would therefore assume that ATA is generally weak compared to niche differentiation (but again, admittedly based on few datapoints). Obviously, the authors may disagree with my personal view, in this case I’d like to see a discussion of this. \\

@@

Note also that Chu and Adler (2015) say "we now believe that the models we fit may not be appropriate for rigorously quantifying fluctuation-dependent mechanisms. For example, the temporal storage effect requires positive covariance between the effects of competition and the environment (Chesson and Huntly 1989): in years most favorable to a species, it should also be the most vulnerable to competition. However, we did not specify our vital rate regressions in a way that makes the covariation between the environment (or random year effects) and competition explicit. As a result, our model may not allow full expression of a temporal storage effect and the simulations may underestimate its strength. Such differences in model specification may explain why a previous analysis, based on a very different modeling approach, showed a strong storage effect in the Kansas community (Adler et al. 2006) in contrast to the weak effect we observed here. We plan to address this model specification issue in future work focused on the population and community consequences of environmental variability."

Note that the Stump and Vasseur paper very clearly says that they don't think storage effects are important. 
It's a 73-page ms, I don't have time to absorb it. This will be part of acknowledging the viewpoint. Maybe Jasmin
has time to read it and decide how it relates to our paper? Do the assumptions they explore also apply
to the Ellner approach of quantifying storage effects?

Finally, I notice that sometimes for our results, ATA effects and effects of correlation per se counter
each other. If that happens, storage effects could be weak, while ATA effects are still strong.

I suppose one must always accept the possibility that future work will render one's results less
meaningful or obsolete.
>} \\

\noindent C12) I like the conceptual idea of removing the long tail. \\

@@

\noindent ***\emph{Thank you for the positive feedback.} \\

\noindent C13) However, I have somewhat difficulties to precisely understand/be sure I correctly understood what this does. I suggest you move a simpler version of Fig S1 into the main text and show the actual distribution, the uncorrelated distribution of $E$ and $C$, but same marginal distribution, and the distribution with the same covariance, but the ATA removed. Additionally, it might be beneficial to indicate which distribution is then used to compute which part of the decomposition. For example, in Figure S1 d you show $E_i^{||}$, but from my understanding this distribution would be used to compute $E_i^{(EC)}$, correct? \\

@@

\noindent ***\emph{<Clearly since two referees have now mentioned it we need to put some version of Fig S1 back into the
main text. The earlier referee wanted to see the whole figure brought back, but this referee wants to see just part, i.e.,
a simpler version, as specified. It's not clear to me how it would help the reader to just see the three panels the referee
suggests. Only 2 of the suggested panels are currently part of Fig. S1 - the suggestion of putting in the uncorrelated
distribution would be new. The current Fig. S1 serves the purpose of describing exactly the process of removing ATAs. 
The new suggested figure would not serve that purpose, it would instead maybe give the reader a sense for what these
distributions look like when manipulated in the ways we are talking about. But it's not real clear to me that much
intuition will be conveyed by doing that. One way to reply would be just to add a panel to the figure that shows 
the uncorrelated case, and move the resulting 5-panel figure back in. Then we'd explain that 1) the other referee asked
for the whole figure, 2) it's not clear to us that the simplified version serves any good purpose, it certainly 
does not serve the purpose of explaining the process of removing ATAs, but 3) we went ahead and added the uncorrelated
case anyway, since that's what was used previously, so it makes sense to have that point of comparison. If we do
this, Jasmin can add the new panel to the figure.>} \\

\noindent C14) You focus solely on storage effect with the argument that it’s the only where ATA can actually occur. But on a broader level you essentially decompose the EC distribution into it’s different momentums. One only carrying about the mean (i.e. first order correct), one with only carrying about the mean and covariance (i.e. getting the first two orders correct), and one with all other orders correct. Given this interpretation of your work, I might apply the same logic to the relative-nonlinearity, where I decompose the relative non-linearity into the variance term and the term with long tails. (one might also even decompose the actual distribution into more distributions, getting successively more and more momentums of the distribution correct, but I would argue that this is probably less useful). \\

\noindent ***\emph{We are not sure we understand precisely what the referee is saying here because their language
appears more suggestive than formally correct. Perhaps they mean
``moment'' rather than ``momentum''? Perhaps they mean ``caring'' instead of ``carrying''? I think the phrase 
``term with long tails'' does not precisely relate to what we're doing since marginal distributions
of the bivariate distribution $(E,C)$ are unchanged in all our approaches, and hence tails are unchanged. 
It's just relationships
between $E$ and $C$ that are modified in the different scenarios we consider. The referee's ideas, here,
if made precise, may be about potentially using a series of approximations to the bivariate distribution $(E,C)$
that have the same marginals as $(E,C)$ itself, but that get successively more co-moments right, compared to the
co-moments of $(E,C)$ itself (and the 
remaining co-moments would be zero, we presume, if
such distributions can be defined). Such an approach, applied to relative nonlinearity, would not give
anything interesting because relative nonlinearity only depends on the marginal distributions, which, 
under this framework, are always equal to the original marginals of $(E,C)$. On the other hand,
perhaps the referee means to take a series of approximations of $(E,C)$ so that the $n$th 
approximate distribution gets both moments and co-moments of order $\leq n$ right, with remaining
moments and co-moments zero (if such distributions can be defined). That may be interesting, but would
certainly be a different paper from ours, since our paper is about \emph{relationships} between $E$ and $C$
(while leaving the univariate distributions of $E$ and $C$ themselves fixed), whereas this idea of the 
referee would make approximations
both to the relationship between $E$ and $C$ and to the marginals, simultaneously. Nevertheless, 
we are pleased the referee 
is interested in our approach to the extent that they are generating potential ideas for future research.} \\

\noindent C15) L34: I found this sentence confusing due to the two “betweens” and it’s not immediately clear to what they “bound” \\

\noindent ***\emph{We modified the sentence.} \\

\noindent C16) L41: The “we” is slightly confusing because the author teams share only one author \\

\noindent ***\emph{We changed the wording here.} \\

\noindent C17) L48: I do not know these on the top of my head, and likely many other readers will neither, so maybe quickly refresh our memory \\

\noindent ***\emph{There is no space to intro these things, so we limited this text to just mentioning 
extinction risks and stability of ecosystem functioning. Whereas some ecologists may not have heard of 
Taylor's law, we feel confident that extinction risks and stability of ecosystem functioning will ring
a bell for most ecologists.} \\

\noindent C18) L49: This is not immediately clear to me why (obviously, you’ve shown this in the other paper). Is it possible to briefly state why? \\

\noindent ***\emph{We spent 15-20 minutes or so trying to add an explanation that was succinct enough to not break
the flow of the paper while still being informative, but we could not do it. Any more than a short sentence would break the 
flow, and we felt that all the one-sentence explanations we could come up with were unclear. 
FYI, in a nutshell the idea is as follows. Assume that low values of an environmental variable reduce the 
growth rates of local populations and high values elevate those growth rates. In that case one may
expect that left-tail associations increase metapopulation extinction risks because then environmental
``catastrophes'' are spatially synchronized, presumably reducing the potential for rescue effects. Whereas
``bonanzas'' (extremely good conditions, with a lot of growth) are much less correlated. The case where
high values of the variable reduce the growth and low values elevate it is similar, but it's right-tail association
that elevates extinction risk. This is most of the explanation, but it gets a bit more nuanced, still - 
see the paper for additional details. Thanks for your interest, by the way.} \\

\noindent C19) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\

\noindent ***\emph{It is not necessary for the reader to understand what these are at this point, and,
in fact, the reader does not need to understand what relative nonlinearity is at all for the main text
of the paper. We apologize for giving the false impression, here, that more background knowledge was needed
than is actually the case. We eliminated the reference to relative nonlinearity and indicated to the 
reader that storage effects would be defined below, which they are. } \\

\noindent C20) Sidenote: After writing these few comments it appears to me that you seem to have a very knowledgeable reader in mind, which may not be the case for a general journal as Ecology letters. \\

\noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce a clear paper, but, of the comments 
C16-C19 above, only C19 relates at all to whether this paper is written for a general ecological audience,
and we have revised the text C19 mentions. 
For instance, the manuscript text mentioned in C17-C18 was just linking this paper to other prior work on ATAs. 
It's not necessary for the reader to understand that work to understand this paper, the point is just to make 
the link to that other literature and to show that ATAs have been found to be important in ecology, generally.
Furthermore, we don't think our paper is any more specialized than the papers Ellner et al (2016, 2019) which
we cite, and which inspired our paper, and which were published in Ecology Letters. 
Nevertheless, we have gone through the paper and redoubled our efforts to make the paper broadly accessible.} \\
%We have to go through the paper and redoubled our efforts to make the paper broadly accessible.
